---
title: "Upscaling: Forest Case Study"
author: "Laura Graham"
output: 
  html_document:
    toc: true
    keep_md: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)
```

```{r setup}
library(MASS)
library(winmoveR)
library(raster)
library(rgdal)
library(rgeos)
library(cowplot)
library(sf)
library(knitr)
library(GGally)
library(e1071) # optimising transformations (skewness function)
library(broom)
library(MuMIn)
library(stringr)
library(DHARMa)
library(tidyverse)


# set up plotting options
theme_set(theme_classic() + theme(strip.background = element_blank()))

varorder <- c("sprich", "winvar2000", "elevvar", "elevmean", "bio1", "bio12", "bio15")

varlabel <- c("Species richness", "MW Elevation (2km)", "LS Elevation", "Mean Elevation", "Temperature", "Precipitation", "Precip. seasonality")

coefforder <- c("(Intercept)", "winvar2000", "elevvar", "elevmean", "elevmean:winvar2000", "elevmean:elevvar", "bio1", "bio12", "bio15", "I(bio1^2)", "I(bio12^2)")

coefflabel <- c("(Intercept)", "MW elevation", "LS elevation", "Mean elevation", "Mean elevation : MW elevation", "Mean elevation : LS elevation", "Temperature", "Precipitation", "Precipitation\nseasonality", "Temperature\n(quadratic)", "Precipitation\n(quadratic)")
```

# Data

## EU forest data

First we need to load in and spatialise the EU forest data: available for download from [figshare](https://ndownloader.figshare.com/files/6662535)

```{r euforest_data}
forests <- read_csv("~/DATA/BIOLOGICAL/eu_forests/eu_forest_species.csv") %>% 
  mutate(sp_name = `SPECIES NAME`) %>% 
  dplyr::select(X, Y, sp_name)

forests <- SpatialPointsDataFrame(forests[,1:2], forests[,3], proj4string = crs("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"))
```

## Environmental data

Analyses will be at 0.5 degree resolution. The data to be upscaled are the elevation data from the [European Environment Agency](https://www.eea.europa.eu/data-and-maps/data/eu-dem). By cropping all datasets to the extent of the EU elevation data, we end up losing some of the data from the EU forest dataset, such as the records from Northern Africa and the Canaries. We aggregate the elevation to 100m (from 25m) for ease of computation.

We will use the moving window to upscale the variation in elevation using a radius of 2km. We will also calculate variation in elevation at 0.5 degree resolution. As covariates, we  calculated average temperature and precipitation at 0.5 degree resolution. 

```{r envdata}
# This section only gets run the first time (or if changed) and the output saved
# to file. Set eval = FALSE at all other times.

# get climate data
wc_bio <- getData('worldclim', var = "bio", res = 10, path = "~/DATA/CLIMATE/worldclim/")
wc_bio_30 <- aggregate(wc_bio, fact = 3)

# load in the elevation data
#elev <- raster("~/DATA/PHYSICAL/elev/eu_dem/eudem_dem_3035_europe.tif")
elev <- raster("~/DATA/PHYSICAL/elev/eu_dem/eudem_100m_aggregated.tif")

# get forest and climate data in same projection as elev and crop worldclim data by the EU forest dataset
forests <- spTransform(forests, crs(elev))
wc_bio_30 <- projectRaster(wc_bio_30, crs = crs(elev))

# crop climate data
bounds <- extent(elev) 
wc_bio_eu <- crop(wc_bio_30, bounds)

# rasterise the EU forest data to create total species richness (the input function gets the total number of unique species in a cell)
forests_r <- rasterize(forests, wc_bio_eu, fun=function(x, ...){ length(unique(na.omit(x)))})[[2]]

# create a grid for the moving window and elevation aggregation to loop through - I want to move this into the package at some point
eu_grid <- as(forests_r, 'SpatialPolygonsDataFrame')

# agg_cell <- function(grid, dat, cell_no, fact, fn="mean", ...) {
#   cell <- grid[cell_no,]
#   dat_cell <- crop(dat, cell, filename='temp_raster/temp_crop.tif', overwrite=TRUE)
#   dat_cell_agg <- aggregate(dat_cell, fact, fun = fn, 
#                             filename='temp_raster/temp_agg.tif', overwrite=TRUE)
#   zero_to_na <- function(x) ifelse(x==0, NA, x)
#   dat_cell_na <- raster::calc(dat_cell_agg, zero_to_na, ...)
# }
# 
# elev_cell_100 <- list()
# 
# for(cell_no in 1:length(eu_grid)) {
#   elev_cell_100[[cell_no]] <- agg_cell(eu_grid, elev, cell_no, 4, filename=paste0('temp_raster/temp_agg_', cell_no, '.tif'), overwrite=TRUE)
# }
# 
# # need to add the arguments to mosaic as attributes of the raster list.
# rasters.mosaicargs <- elev_cell_100
# names(rasters.mosaicargs) <- NULL
# rasters.mosaicargs$fun <- mean
# rasters.mosaicargs$filename <- '~/DATA/PHYSICAL/elev/eu_dem/eudem_100m_aggregated.tif'
# rasters.mosaicargs$overwrite <- TRUE
# rasters.mosaicargs$na.rm <- TRUE
# 
# elev_100 <- do.call("mosaic", rasters.mosaicargs)

elev_100 <- elev

# create the grid level covariates
eu_grid$elevvar <- nomove_upscale(grid = eu_grid, dat = elev_100, fn = "var", na.rm = TRUE)

eu_grid$elevmean <- nomove_upscale(grid = eu_grid, dat = elev_100, fn = "mean", na.rm = TRUE)

# upscale the eu elevation data using the moving window approach
strt <- Sys.time()
eu_grid$winvar2000 <- winmove_upscale(eu_grid, elev_100, 2000, 
                                        "rectangle", "var")
run_time <- difftime(Sys.time(), strt, units = "mins")

save(run_time, file = "results/forest_runtime.Rda")

forests_pts <- spTransform(gCentroid(eu_grid, byid=TRUE), crs(elev))

forests_sp <- st_as_sf(eu_grid) %>% 
  bind_cols(raster::extract(wc_bio_eu, forests_pts) %>% as.tibble) %>% 
  rename(sprich = sp_name)

save(forests_sp, file="results/forests_covariates_sp.Rda")
```


`r load("results/forest_runtime.Rda")`
Upscaling 100m resolution elevation data for Europe to 0.5 degrees using a 2km radius window took:  `r run_time` minutes. 

## Exploration and transformation

What do the variables look like spatially?

```{r spatial_plot, fig.width = 10}
load("results/forests_covariates_sp.Rda")
load("results/forests_covariates.Rda")
forests_narrow <- forests_sp %>% 
  select(varorder) %>% 
  mutate_at(.vars = vars(-sprich, -geometry), .funs = funs(scale)) %>% 
  gather(variable, value, -geometry) %>% 
  mutate(facet = "Tree species richness") %>% na.omit

sp_plot <- ggplot(forests_narrow %>% filter(variable == "sprich")) + 
  geom_sf(aes(fill = value), colour = NA) +
  coord_sf(crs = st_crs(forests_narrow), datum = NA) + 
  scale_fill_viridis_c(name = "", option = "magma") + 
  facet_wrap(~facet) + 
  theme(axis.text = element_blank(), axis.line = element_blank(), axis.ticks = element_blank(),
        legend.position = "bottom", legend.title.align = 0.5,
        legend.key.height=unit(6,"points"), legend.key.width = unit(1.5, "line"))

covs <- forests_narrow %>% 
  filter(variable %in% varorder[-1]) %>% 
  mutate(variable = factor(variable, levels = varorder, labels = varlabel))

cov_plot <- ggplot(covs) + 
  geom_sf(aes(fill = value), colour = NA) +
  coord_sf(crs = st_crs(forests_narrow), datum = NA) + 
  scale_fill_viridis_c(name = "") + facet_wrap(~variable) + 
  theme(axis.text = element_blank(), axis.line = element_blank(), axis.ticks = element_blank(),
        legend.position = "bottom", legend.title.align = 0.5,
        legend.key.height=unit(6,"points"), legend.key.width = unit(2, "line"))

spatial_plot <- plot_grid(sp_plot, cov_plot, labels = c("a)", "b)"),
                          label_size = 10, rel_widths = c(1, 1.4))

save(spatial_plot, file = "results/forest_spatial.Rda")

spatial_plot
```

How are the variables distributed and where are the correlations?

```{r pairs, fig.width = 10, fig.height = 10}
forests_df <- forests_sp %>% as.tibble %>% select(varorder)

ggpairs(
  forests_df %>% select(varorder), 
  upper = list(
    continuous = wrap('cor', method = "spearman")
  ),
  columnLabels = varlabel
)
```

MW Elevation is the lowest correlation except for total precipitation (bio12). Highest correlation is with precipitation seasonality (bio15). Our expectation that variation in elevation would be important is supported here by the fact that the species richness correlation with mean elevation is lower. Species richness ideal for Poisson distribution; right skew to MW Elevation and LS elevation, Mean elevation, as well as the precipitation variables - need to transform. 

```{r transform}
# optimise to find optimal value for log transform
skew.score <- function(c, x) (skewness(log(x + c)))^2
c.winvar2000 <- optimise(skew.score, c(0, 20), x = forests_df$winvar2000)$minimum
c.elevvar <- optimise(skew.score, c(0, 20), x = forests_df$elevvar)$minimum
c.elevmean <- optimise(skew.score, c(0, 20), x = forests_df$elevmean)$minimum
c.bio12 <- optimise(skew.score, c(0, 20), x = forests_df$bio12)$minimum
c.bio15 <- optimise(skew.score, c(0, 20), x = forests_df$bio15)$minimum


forests_df_t <- select(forests_df, varorder) %>% 
  mutate(winvar2000 = log(winvar2000 + c.winvar2000),
         elevvar = log(elevvar + c.elevvar),
         elevmean = log(elevmean + c.elevmean),
         bio12 = log(bio12 + c.bio12),
         bio15 = log(bio15 + c.bio15))

forests_narrow_t <- gather(forests_df_t, variable, value) %>% 
  mutate(variable = factor(variable, levels = varorder, labels = varlabel))

ggplot(forests_narrow_t, aes(x = value)) + 
  geom_histogram(bins = 30) + 
  facet_wrap(~variable, scales = "free_x")
```

We scaled the data (mean = 0, sd = 1) so that the partial regression coefficients are comparable. 

```{r scale_data}
scale_this <- function(x) as.vector(scale(x))
forests_df_t <- mutate_at(forests_df_t, .vars = vars(-sprich), .funs = funs(scale_this))
```

# Models

## Global model

We are including the quadratic term for temperature (bio1) and precipitation (bio12), due to the shape of the relationship between these variables (and based on some earlier residual diagnostics)

```{r global_model}
mod_global <- glm.nb(sprich ~ elevmean + winvar2000 + elevvar + elevmean:winvar2000 + elevmean:elevvar + bio1 + I(bio1^2) + bio12 + I(bio12^2) + bio15, data = forests_df_t, na.action = na.fail)

res_global <- mod_global %>% coef %>% enframe(name = "variable", value = "coef") %>%
  left_join(mod_global %>% confint %>% as_tibble(rownames = "variable")) %>%
  mutate(fvariable = factor(variable, levels = coefforder, labels = gsub("\n", " ", coefflabel))) %>%
  arrange(fvariable) 

globalr2 <- round((1 - (mod_global$deviance / mod_global$null.deviance))*100, 2)

res_global %>% select(fvariable, coef, `2.5 %`, `97.5 %`) %>% kable(digits=3)
```

There is a negative effect of local-scale (2km) variation in elevation (MW Elevation), but a positive effect of landscape-scale (~50km) variation in elevation (LS Elevation). 

This model explains `r globalr2`% of the deviance in tree species richness. This was calculated using D-squared. 

Check the model specification using the DHARMa package. 

```{r global_validation, fig.height = 8}
simulationOutput <- simulateResiduals(fittedModel = mod_global, n = 250)
plotSimulatedResiduals(simulationOutput)
plotResiduals(forests_df_t$winvar2000, simulationOutput$scaledResiduals, xlab = "MW elevation (2km)")
plotResiduals(forests_df_t$elevvar, simulationOutput$scaledResiduals, xlab = "LS elevation")
plotResiduals(forests_df_t$elevmean, simulationOutput$scaledResiduals, xlab = "Mean elevation")
plotResiduals(forests_df_t$bio1, simulationOutput$scaledResiduals, xlab = "Temperature")
plotResiduals(forests_df_t$bio12, simulationOutput$scaledResiduals, xlab = "Precipitation")
plotResiduals(forests_df_t$bio15, simulationOutput$scaledResiduals, xlab = "Precipitation seasonality")
```

Based on the residual diagnostics, have gone with a negative binomial model due to overdispersion. The earlier version of the diagnostics also found patterns with temperature (bio1) and precipitation (bio12), hence the inclusion of their quadratic terms.

## Parsimonious model

This model is determined using stepwise selection. 

```{r simple_model, message = TRUE}
mod_simple <- step(mod_global)

res_simple <- mod_simple %>% coef %>% enframe(name = "variable", value = "coef") %>%
  left_join(mod_global %>% confint %>% as_tibble(rownames = "variable")) %>%
  mutate(fvariable = factor(variable, levels = coefforder, labels = gsub("\n", " ", coefflabel))) %>%
  arrange(fvariable)
simplep <- glance(mod_simple)$p.value
simpler2 <- glance(mod_simple)$r.squared

res_simple %>% select(fvariable, coef, `2.5 %`, `97.5 %`) %>% kable(digits=3)
```

Full model retained. 

## Model average in MMI framework

Now to get the model averaged and variable importance estimates. We are using `dredge` to get the full list of models, then `model.avg` to get the estimates for the 95% confidence set. 

```{r mmi_model}
mod_mmi <- dredge(mod_global, beta = "none") %>% 
  model.avg(subset = cumsum(weight) <= 0.95)

# mmi averaged estimates
res_mmi <- mod_mmi %>% coef %>% enframe(name = "variable", value = "coef") %>%
  left_join(mod_mmi %>% confint %>% as_tibble(rownames = "variable")) %>%
  left_join(mod_mmi %>% importance %>% enframe(name = "variable", value = "importance")) %>% 
  mutate(fvariable = factor(variable, levels = coefforder, labels = gsub("\n", " ", coefflabel))) %>%
  arrange(fvariable)

res_mmi %>% select(fvariable, coef, `2.5 %`, `97.5 %`, importance) %>% kable(digits = 3, caption = "Results of model averaging")
```

# Plots and summary of main results

The window-based elevation measure was included in all supported models (summed Akaike weight = `r res_mmi %>% filter(variable == "winvar2000") %>% select(importance) %>% round(2)`, as was the landscape elevation measure (summed Akaike weight = `r res_mmi %>% filter(variable == "elevvar") %>% select(importance) %>% round(2)`). MW elevation and LS elevation had a similar absolute effect size based on the partial regression coefficients. However, MW elevation had a negative effect on species richness, whereas for LS elevation it was positive. This suggests that topographic variation has different effects on tree species richness depending on the scale at which it is measured. The interaction between mean elevation and MW elevation was positive, and present in all supported models. This suggests that effect of local-scale topographic variation is positive at high altitudes, but negative at low altitudes. The opposite is true of LS elevation, although this interaction term is not as strong, and was not present in all supported models (summed Akaike weight = `r res_mmi %>% filter(variable == "elevmean:elevvar") %>% select(importance) %>% round(2)`)

```{r results_out, fig.width = 15, fig.height = 8}
# This function lets us calculate the coefficient estimate and 95% CI for one
# variable at a particular value of the other variable in an interaction
calc_estci <- function(mod, variable1, variable2, value) {
  mod <- get(mod)
  slope <- coef(mod)[variable1]
  # catch it so either order will work
  int_slope <- coef(mod)[paste0(variable1, ":", variable2)]
  if(is.na(int_slope)) {
    int_slope <- coef(mod)[paste0(variable2, ":", variable1)]
  }
  var1 <- vcov(mod)[variable1, variable1]
  var2 <- vcov(mod)[variable2, variable2]
  cov_int <- vcov(mod)[variable1, variable2]
  UCI <- (slope + int_slope * value) + (1.9602 * sqrt(var1 + (2 * value * cov_int) + ((value^2) * var2)))
  LCI <- (slope + int_slope * value) - (1.9602 * sqrt(var1 + (2 * value * cov_int) + ((value^2) * var2)))
  y <- (slope + int_slope * value)
  return(tibble(coef = y, `2.5 %` = LCI, `97.5 %` = UCI))
}

# The first line creates a dataframe of the variables to calculate interaction for
figure1_df <- tibble(mod = "mod_global", 
                     variable1 = c(rep("winvar2000", 2), rep("elevvar", 2)),
                     variable2 = "elevmean", value = c(-1, 1, -1, 1)) %>%
  bind_cols(pmap_df(., calc_estci)) %>%
  inner_join(tibble(value = c(-1, 1), type = c("low", "high"))) %>%
  select(variable = variable1, type, coef, `2.5 %`, `97.5 %`) %>% 
  bind_rows(res_global %>% mutate(type = "mean") %>% select(-fvariable)) %>%
  filter(variable != "(Intercept)") %>%
  filter(!str_detect(variable, "elevmean:")) %>%
  mutate(variable = factor(variable, levels = coefforder, labels = coefflabel),
         type = factor(type, levels = c("low", "mean", "high"))) %>%
  na.omit

pd <- position_dodge(0.5)

forest_figure <- ggplot(figure1_df, aes(x = variable, y = coef, colour = type)) + 
  geom_errorbar(aes(ymin=`2.5 %`, ymax=`97.5 %`, colour = type), width=0.2, position=pd) +
  geom_point(position = pd) +
  scale_color_manual(values = c("grey", "black", "grey")) +
  geom_hline(yintercept = 0, linetype = "dotted") + 
  ylab(expression(paste("Coefficient estimate " %+-% " 95% CI"))) + 
  xlab("") + theme(legend.position = "none")

forest_figure

save(forest_figure, file = "results/forest_res_figure.Rda")
```