---
title: "Upscaling: Forest Case Study"
author: "Laura Graham"
output: 
  html_document:
    toc: true
bibliography: C:\Users\lg1u16\Documents\mendeley_bibtex\library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
library(winmoveR)
library(raster)
library(rgdal)
library(rgeos)
library(ggmap)
library(cowplot)
library(sf)
library(tidyverse)
```

# Case study 2: Forest 

## Data

### EU forest data

First we need to load in and spatialise the EU forest data. These are data from @Mauri2017. Available for dowload on [figshare](https://ndownloader.figshare.com/files/6662535)
```{r euforest_data}
forests <- read_csv("../DATA/eu_forests/eu_forest_species.csv") %>% 
  mutate(sp_name = `SPECIES NAME`) %>% 
  dplyr::select(X, Y, sp_name)

forests <- SpatialPointsDataFrame(forests[,1:2], forests[,3], proj4string = crs("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"))
```

### Environmental data

The data to be upscaled are the elevation data from the [European Environment Agency](https://www.eea.europa.eu/data-and-maps/data/eu-dem). Note that these are in a specific European projection: European Grid: ETRS89-LAEA. All other datasets will be transformed to this projection. By cropping all datasets to the extent of the EU elevation data, we end up losing some of the data from the EU forest dataset, such as the records from Northern Africa and the Canaries. We aggregate the elevation to 100m (from 25m) both for ease of computation and *is there a biological reason??* - also characteristic scale of variation? doesn't really change visually much from 25m to 100m - need to do some kind of test to make sure this simplification if valid. 

We will use the moving window to upscale the variation in elevation at three scales: 500 m, 1 km, 2 km. We do not have *a priori* expectation of the appropriate scale of effect, but we do have reason to expect that variation in elevation will determine species richness due to the opening up of niche space [citations]. We will also calculate variation in elevation at the 0.5 degree scale. As covariates, to avoid confounding variables, we will calculate average temperature and precipitation at the 0.5 degree scale. 

```{r envdata, eval = FALSE}
# This section only gets run the first time (or if changed) and the output saved
# to file. Set eval = FALSE at all other times.

wc_bio <- getData('worldclim', var = "bio", res = 10, path = "../data/worldclim/")
wc_bio_30 <- aggregate(wc_bio, 3)

# load in the elevation data
elev <- raster("../../data/elev/eudem_dem_3035_europe.tif")

# get forest and climate data in same projection as elev and crop worldclim data by the EU forest dataset
eu_forests_pts <- spTransform(eu_forests_pts, crs(elev))
wc_bio_30 <- projectRaster(wc_bio_30, crs = crs(elev))

# crop climate data
bounds <- extent(elev) 
wc_bio_eu <- crop(wc_bio_30, bounds)

# rasterise the EU forest data to create total species richness (the input function gets the total number of unique species in a cell)
eu_forests_r <- rasterize(eu_forests_pts, wc_bio_eu, fun=function(x, ...){ length(unique(na.omit(x)))})[[2]]

# create a grid for the moving window and elevation aggregation to loop through - I want to move this into the package at some point
eu_grid <- as(eu_forests_r, 'SpatialPolygonsDataFrame')

agg_cell <- function(grid, dat, cell_no, fact, fn="mean", ...) {
  cell <- grid[cell_no,]
  dat_cell <- crop(dat, cell, filename='../temp_raster/temp_crop.tif', overwrite=TRUE)
  dat_cell_agg <- aggregate(dat_cell, fact, fun = fn, filename='../temp_raster/temp_agg.tif', overwrite=TRUE)
  zero_to_na <- function(x) ifelse(x==0, NA, x)
  dat_cell_na <- calc(dat_cell_agg, zero_to_na, ...)
}

elev_cell_100 <- list()

for(cell_no in 1:length(eu_grid)) {
  elev_cell_100[[cell_no]] <- agg_cell(eu_grid, elev, cell_no, 4, filename=paste0('../temp_raster/temp_agg_', cell_no, '.tif'), overwrite=TRUE)
}

# need to add the arguments to mosaic as attributes of the raster list. 
rasters.mosaicargs <- elev_cell_100
rasters.mosaicargs$fun <- mean
rasters.mosaicargs$filename <- '../../data/elev/eudem_100m_aggregated.tif'
rasters.mosaicargs$overwrite <- TRUE
elev_100 <- do.call(mosaic, rasters.mosaicargs)

# create the grid level elevation variance
elev_grid <- aggregate(elev_100, res(eu_forests_r)/res(elev_100), fun = var,
                       filename='../temp_raster/elev_grid.tif', overwrite=TRUE)

# upscale the eu elevation data using the moving window approach
eu_grid$win_var_500 <- winmove_upscale(eu_grid, elev_100, 500, "rectangle", "var", na.rm=TRUE)$V1
eu_grid$win_var_1000 <- winmove_upscale(eu_grid, elev_100, 1000, "rectangle", "var", na.rm=TRUE)$V1
eu_grid$win_var_2000 <- winmove_upscale(eu_grid, elev_100, 2000, "rectangle", "var", na.rm=TRUE)$V1

# stack variables
win_var_500 <- rasterize(eu_grid, eu_forests_r, field = eu_grid@data$win_var_500, fun = "mean")
win_var_1000 <- rasterize(eu_grid, eu_forests_r, field = eu_grid@data$win_var_1000, fun = "mean")
win_var_2000 <- rasterize(eu_grid, eu_forests_r, field = eu_grid@data$win_var_2000, fun = "mean")

var_list <- lapply(list(eu_forests_r, win_var_500, win_var_1000, win_var_2000, elev_grid, wc_bio_eu),
                   function(x) crop(x, elev_grid))
var_stack <- stack(var_list)
names(var_stack)[1:5] <- c("sp_rich", "win_var_500", "win_var_1000", "win_var_2000", "elev_var")
writeRaster(var_stack, filename="../results/eu_forests_vars.tif", options="INTERLEAVE=BAND", overwrite=TRUE)
var_df <- as.data.frame(var_stack) %>% na.omit %>% 
  write_csv("../results/eu_forests_vars.csv")
```

Now all of the variables are in one place, let's do some analysis on them:

```{r cor}
var_df_temp <- dplyr::select(var_df, sp_rich, win_var_500, win_var_1000, win_var_2000, elev_var, bio1, bio12)
plot(var_df)
```
