---
title: "Appendix III: *Garrulus glandarius case study"
output: pdf_document
header-includes:
    - \usepackage{booktabs}
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Required packages

In the first step, we load all of the required packages. Note that winmoveR (the package developed for this paper) can be installed using `devtools::install_github("laurajanegraham/winmoveR")`

The development version of ggplot2 is required for the spatial plots. Install using `devtools::install_github("tidyverse/ggplot2")`

Other options in this section set the order and labels for the variables for plotting and tables. 

```{r setup}
library(winmoveR)
library(raster)
library(rgdal)
library(rgeos)
library(sf)
library(knitr)
library(GGally)
library(e1071) # optimising transformations (skewness function)
library(broom)
library(MuMIn)
library(stringr)
library(DHARMa)
library(cowplot)
library(tidyverse)


# set up plotting options
theme_set(theme_classic() + theme(strip.background = element_blank()))

# set up some options
# for exploratory plotting
varorder <- c("index10", "winshannon", "lsshannon", "habitat", "urban", "bio1")

varlabel <- c("Abundance (2010 Atlas)", "MW Shannon", "LS Shannon", 
              "Forest %", "Urban %", "Temperature")

# for initial tables from models
coefforder <- c("(Intercept)", "winshannon", "lsshannon", "habitat", 
               "habitat:winshannon", "habitat:lsshannon",
               "urban", "bio1")

coefflabel <- c("Intercept", "MW Shannon", "LS Shannon", "Forest %", 
                    "Forest % : MW Shannon", "Forest % : LS Shannon", 
                    "Urban %", "Temperature")
```

# Data

## Relative abundance index

First we need to load in and spatialise the Jay (*Garrulus glandarius*) data. These were provided by Simon Gillings at the BTO. The data includes the relative abundance indices for the 1990 Atlas (index90) and for the 2010 Atlas (index10). We are using the 2010 Atlas data. 

```{r bto_data, eval = FALSE}
jay_sp <- read_sf("~/DATA/ADMINISTRATIVE/bng/10km_grid_region.shp") %>% 
  rename(grid = TILE_NAME) %>% 
  inner_join(read_csv("~/DATA/BIOLOGICAL/bto_jays/extract.csv") %>% 
               rename(grid = tenkm)) %>%
  select(index10, geometry) %>%
  filter(!is.na(geometry))
```

## Environmental data

The data for which we want to use the upscaling approach on is Land Cover Map 2007, which is the closest match to the 2010 relative abundance index for jays. We will use the moving window to upscale diversity of the two used habitats: Broadleaf and Coniferous forest (LCM codes 1, 2). Eurasian jays use a combination of these habitats: broadleaf for foraging, coniferous for nesting. We will calculate Shannon Diversity on just these two habitats to create a measure of the landscape structure used by jays. We will calculate Shannon diversity using the moving window approach at the 1km scale [needs citation] and without the moving window at the 10km scale. As covariates, we will calculate forest (habitat) cover percentage, urban land cover percentage and from Worldclim the following bioclimatic variables at the 10km scale: temperature (bio1), temperature range (bio7), annual preciptation (bio12) and precipitation seasonality (bio15). 

```{r envdata, eval = FALSE}
# This section only gets run the first time (or if changed) and the output saved to file.
# Set eval = FALSE at all other times.
lcm <- raster("~/DATA/LULC/lcm2007/lcm2007_25m_gb.tif")
forest <- c(1, 2) # these are the two forest classes
urban <- c(22, 23)

# grid needs to be SpatialPolygonsDataFrame until I update the winmoveR package to use sf
# objects
bng <- as(jay_sp, 'Spatial')

# shannon via moving window
strt <- Sys.time()
bng$winshannon <- winmoveR::winmove_upscale(grid = bng, dat = lcm, radius = 1000, 
                                            type = "rectangle", fn = "diversity", 
                                            lc_class = forest)
run_time <- difftime(Sys.time(), strt, units = "mins")

save(run_time, file = "results/jays_runtime.Rda")

# shannon without moving window
bng$lsshannon <- nomove_upscale(grid = bng, dat = lcm, 
                                fn = "diversity", lc_class = forest)

# other measures from lcm
bng$habitat <- nomove_upscale(grid = bng, dat = lcm, fn = "prop", lc_class = forest)
bng$urban <- nomove_upscale(grid = bng, dat = lcm, fn = "prop", lc_class = urban)

# Bioclimatic variables from worldclim
wc_bio <- getData('worldclim', var = 'bio', path = '~/DATA/CLIMATE/worldclim/', res=5)
bng_pts <- spTransform(gCentroid(bng, byid=TRUE), crs(wc_bio))
jay_sp <- st_as_sf(bng) %>% bind_cols(raster::extract(wc_bio, bng_pts) %>% as.tibble)

save(jay_sp, file="results/jays_covariates.Rda")
```

`r load("results/jays_runtime.Rda")`
Upscaling 25m resolution land-cover data for Europe to 10km degrees resolution using a 1km radius window took:  `r round(run_time, 2)` minutes. 

## Exploration and transformation

We need to first remove cells with 0 abundance - these cause issues with the model fit and we are only really interested in predicting abundance presuming the species is present. 

What do the variables look like spatially?

```{r spatial_plot, fig.width = 10}
load("results/jays_covariates.Rda")
jay_sp <- filter(jay_sp, index10 != 0)

jay_narrow <- jay_sp %>% 
  mutate_at(.vars = vars(-index10, -geometry), .funs = funs(scale)) %>% 
  gather(variable, value, -geometry) %>% 
  mutate(facet = "Relative abundance index")

jay_plot <- ggplot(jay_narrow %>% filter(variable == "index10")) + 
  geom_sf(aes(fill = value), colour = NA) +
  coord_sf(crs = st_crs(jay_narrow), datum = NA) + 
  scale_fill_viridis_c(name = "", option = "magma") + 
  facet_wrap(~facet) + 
  theme(axis.text = element_blank(), axis.line = element_blank(), 
        axis.ticks = element_blank(),
        legend.position = "bottom", legend.title.align = 0.5,
        legend.key.height=unit(6,"points"), legend.key.width = unit(1.5, "line"))

covs <- jay_narrow %>% 
  filter(variable %in% varorder[-1]) %>% 
  mutate(variable = factor(variable, levels = varorder, labels = varlabel))

cov_plot <- ggplot(covs) + 
  geom_sf(aes(fill = value), colour = NA) +
  coord_sf(crs = st_crs(jay_narrow), datum = NA) + 
  scale_fill_viridis_c(name = "") + facet_wrap(~variable) + 
  theme(axis.text = element_blank(), axis.line = element_blank(), 
        axis.ticks = element_blank(),
        legend.position = "bottom", legend.title.align = 0.5,
        legend.key.height=unit(6,"points"), legend.key.width = unit(2, "line"))
  
spatial_plot <- plot_grid(jay_plot, cov_plot, labels = c("a)", "b)"), 
                          label_size = 10, rel_widths = c(1, 1.5))

save(spatial_plot, file = "results/jays_spatial.Rda")

spatial_plot
```

How are the variables distributed and where are the correlations?

```{r pairs}
jay_df <- jay_sp %>% as.tibble %>% select(varorder)

ggpairs(
  jay_df %>% select(varorder), 
  upper = list(
    continuous = wrap('cor', method = "spearman")
  ),
  columnLabels = varlabel
)
```

MW Shannon a stronger correlate of relative abundance than LS Shannon. Although both Shannon measures correlate with the amount of habitat, MW Shannon is least correlated with this (still very high). Right skew to habitat percentage, urban percentage and precipitation, and left skew to temperature. 

Based on distributions (and some initial model diagnostics not shown) need to transform the temperature, habitat and urban percentage variables. Additionally, because the abundance index is proportional, we transform this using a logit transform with the smallest non-zero value added to the numerator and denominator due to presence of 0 and 1 in data:

```{r transform}
# optimise to find optimal value for log transform
skew.score <- function(c, x) (skewness(log(x + c)))^2
c.habitat <- optimise(skew.score, c(0, 20), x = jay_df$habitat)$minimum
c.urban <- optimise(skew.score, c(0, 20), x = jay_df$urban)$minimum
x <- min(jay_df$index10[jay_df$index10 > 0])
jay_df_t <- mutate(jay_df, 
                   index10_logit = log((index10 + x)/(1 - index10 + x)),
                   habitat = log(habitat + c.habitat),
                   urban = log(urban + c.urban),
                   bio1 = bio1^3)

jay_narrow_t <- gather(jay_df_t, variable, value) %>% 
  mutate(variable = factor(variable, 
         levels = varorder, labels = varlabel))

ggplot(jay_narrow_t, aes(x = value)) + 
  geom_histogram(bins = 30) + 
  facet_wrap(~variable, scales = "free_x")

```

The abundance score for two cells was equal to 1, so the smallest non-zero percentage response (`r round(x, 2)`) was added to the logit function to avoid divide by zero issues.

Now scale the data (mean = 0, sd = 1) so that the partial regression coefficients are comparable. 

```{r scale_data}
scale_this <- function(x) as.vector(scale(x))
jay_df_t <- mutate_at(jay_df_t, .vars = vars(-index10), .funs = funs(scale_this))
```

# Models 

For each model, our sample size is n = `r nrow(jay_df_t)`. 

## Global model

This is the model which contains all predictors (window- and landscape-based Shannon measure, habitat amount, and the interaction between habitat amount and the two Shannon measures) and the additional covariates (urban cover and average temperature). 

```{r global_model}
mod_global <- lm(index10_logit ~ habitat + winshannon + lsshannon + urban + bio1 +
                   habitat:lsshannon + habitat:winshannon, 
                 data = jay_df_t, na.action = na.fail)

res_global <- mod_global %>% coef %>% enframe(name = "variable", value = "coef") %>%
  left_join(mod_global %>% confint %>% as_tibble(rownames = "variable")) %>%
  mutate(fvariable = factor(variable, levels = coefforder, labels = coefflabel)) %>%
  arrange(fvariable) 

# check this calculation of R2
globalr2 = glance(mod_global)$r.square

res_global %>% select(fvariable, coef, `2.5 %`, `97.5 %`) %>% kable(digits=3)
```

First model looks to fit expectations - partial regression coefficient for window-based Shannon is larger than the landscape-based Shannon coefficient, habitat (marginally) and the interaction terms. 

The model explains `r round(globalr2*100, 2)`% of the variance in jay abundance. 

Check the model specification:

```{r global_validation, fig.height = 8}
jay_df_t$resids <- mod_global$residuals
jay_df_t$fitted <- mod_global$fitted.values

ggplot(jay_df_t, aes(x = fitted, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0)
ggplot(jay_df_t, aes(x = winshannon, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0) + xlab("MW Shannon")
ggplot(jay_df_t, aes(x = lsshannon, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0) + xlab("LS Shannon")
ggplot(jay_df_t, aes(x = habitat, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0) + xlab("Forest %")
ggplot(jay_df_t, aes(x = urban, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0) + xlab("Urban %")
ggplot(jay_df_t, aes(x = bio1, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0) + xlab("Temperature")

fitted <- exp(jay_df_t$fitted + x)/(1+exp(jay_df_t$fitted + x))
```

There is still some patterning in the residuals, but overall a reasonable model fit. The lowest fitted value is `r round(min(fitted), 2)` and the highest is `r round(max(fitted), 2)`. Much better conformity to assumptions and range of predicted values than either the normal or binomial versions. 

## Parsimonious model

This model is determined using stepwise selection. 

```{r simple_model, message = TRUE}
mod_simple <- step(mod_global)

res_simple <- mod_simple %>% coef %>% enframe(name = "variable", value = "coef") %>%
  left_join(mod_global %>% confint %>% as_tibble(rownames = "variable")) %>%
  mutate(fvariable = factor(variable, levels = coefforder, labels = coefflabel)) %>%
  arrange(fvariable)
simplep <- glance(mod_simple)$p.value
simpler2 <- glance(mod_simple)$r.squared

res_simple %>% select(fvariable, coef, `2.5 %`, `97.5 %`) %>% kable(digits = 3)
```

This model loses the interaction between Forest % and MW Shannon. Conclusions remain the same. This model explains `r round(simpler2*100, 2)`% of the variance. 

Validation of this model: 

```{r simple_validation}
jay_df_t$resids <- mod_simple$residuals
jay_df_t$fitted <- mod_simple$fitted.values

ggplot(jay_df_t, aes(x = fitted, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0)
ggplot(jay_df_t, aes(x = winshannon, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0) + xlab("MW Shannon")
ggplot(jay_df_t, aes(x = lsshannon, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0) + xlab("LS Shannon")
ggplot(jay_df_t, aes(x = habitat, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0) + xlab("Forest %")
ggplot(jay_df_t, aes(x = urban, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0) + xlab("Urban %")
ggplot(jay_df_t, aes(x = bio1, y = resids)) + 
  geom_point() + geom_hline(yintercept = 0) + xlab("Temperature")

fitted <- exp(jay_df_t$fitted + x)/(1+exp(jay_df_t$fitted + x))
```

Residuals are better again. Lowest fitted value is `r round(min(fitted), 2)` and the highest is `r round(max(fitted), 2)`.

## Model average in MMI framework

Now to get the model averaged and variable importance estimates. We are using `dredge` to get the full list of models, then model.avg to get the estimates for the 95% confidence set. 

```{r mmi_model}
mod_mmi <- dredge(mod_global, beta = "none") %>% 
  model.avg(subset = cumsum(weight) <= 0.95)

# mmi averaged estimates
res_mmi <- mod_mmi %>% coef %>% enframe(name = "variable", value = "coef") %>%
  left_join(mod_mmi %>% confint %>% as_tibble(rownames = "variable")) %>%
  left_join(mod_mmi %>% importance %>% 
              enframe(name = "variable", value = "importance")) %>% 
  mutate(fvariable = factor(variable, levels = coefforder, labels = coefflabel)) %>%
  arrange(fvariable)

res_mmi %>% select(fvariable, coef, `2.5 %`, `97.5 %`, importance) %>% 
  kable(digits = 2)
```

MW Shannon measure again coming out as more important than LS Shannon using both the model averaged partial regression coefficients and the sum of Akaike weights (importance in the above table). 

# Plots and summary of main results

MW Shannon was included in more supported models (summed Akaike weight = `r res_mmi %>% filter(variable == "winshannon") %>% select(importance) %>% round(2)` than the LS Shannon (summed Akaike weight = `r res_mmi %>% filter(variable == "lsshannon") %>% select(importance) %>% round(2)`). The greater importance of the window-based measure is also demonstrated by its larger absolute effect and smaller confidence intervals in the global and mmi modelling approaches. LS Shannon was not present in the parsimonious modelling approach. Results for the global model are shown. 

```{r results_out}
# This function lets us calculate the coefficient estimate and 95% CI for one variable at
# a particular value of the other variable in an interaction
calc_estci <- function(mod_name, variable1, variable2, value) {
  mod <- get(mod_name)
  if (paste0(variable1, ":", variable2) %in% names(mod$coefficients) | 
      paste0(variable2, ":", variable1) %in% names(mod$coefficients)) {
    slope <- coef(mod)[variable1]
    # catch it so either order will work
    int_slope <- coef(mod)[paste0(variable1, ":", variable2)]
    if(is.na(int_slope)) {
      int_slope <- coef(mod)[paste0(variable2, ":", variable1)]
    }
    var1 <- vcov(mod)[variable1, variable1]
    var2 <- vcov(mod)[variable2, variable2]
    cov_int <- vcov(mod)[variable1, variable2]
    UCI <- (slope + int_slope * value) + 
      (1.9602 * sqrt(var1 + (2 * value * cov_int) + ((value^2) * var2)))
    LCI <- (slope + int_slope * value) - 
      (1.9602 * sqrt(var1 + (2 * value * cov_int) + ((value^2) * var2)))
    y <- (slope + int_slope * value)
    return(tibble(coef = y, `2.5 %` = LCI, `97.5 %` = UCI))
  } else {
    return(tibble(coef = 0, `2.5 %` = 0, `97.5 %` = 0))
  }
}

# The first line creates a dataframe of the variables to calculate interaction for
figure1_df <- tibble(mod = "mod_global", 
                     variable1 = c(rep("winshannon", 2), rep("lsshannon", 2)),
                     variable2 = "habitat", value = c(-1, 1, -1, 1)) %>%
  bind_cols(pmap_df(., calc_estci)) %>%
  inner_join(tibble(value = c(-1, 1), type = c("low", "high"))) %>%
  select(variable = variable1, type, coef, `2.5 %`, `97.5 %`) %>% 
  bind_rows(res_global %>% mutate(type = "mean") %>% select(-fvariable)) %>%
  filter(variable != "(Intercept)") %>%
  filter(!str_detect(variable, "habitat:")) %>%
  mutate(variable = factor(variable, levels = coefforder, labels = coefflabel),
         type = factor(type, levels = c("low", "mean", "high"))) %>%
  na.omit

pd <- position_dodge(0.5)

jay_figure <- ggplot(figure1_df, aes(x = variable, y = coef, colour = type)) + 
  geom_errorbar(aes(ymin=`2.5 %`, ymax=`97.5 %`, colour = type), width=0.2, position=pd) +
  geom_point(position = pd) +
  scale_color_manual(values = c("grey", "black", "grey")) +
  geom_hline(yintercept = 0, linetype = "dotted") + 
  ylab(expression(paste("Coefficient estimate " %+-% " 95% CI"))) + 
  xlab("") + theme(legend.position = "none")

jay_figure

save(jay_figure, file = "results/jays_res_figure.Rda")
```

## Session Info
```{r}
devtools::session_info()
```