---
title: "Moving window data aggregation"
author: "Laura Graham"
date: "16/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.path = 'figures/',
                      dpi = 300,
                      dev = c('png'),
                      fig.width = 4.33071,
                      fig.height = 4.33071) 

library(tidyverse)
library(NLMR)
library(raster)
library(cowplot)
library(grainchanger)
library(sf)

# set up plotting options
theme_set(theme_bw(base_size = 7) + theme(strip.background = element_blank(), 
                             panel.grid.major = element_blank(),
                             panel.grid.minor = element_blank(),
                             plot.title = element_text(face = "bold")))

```

# Moving window approach

This is the code to generate figure 1 from the paper. 

## Initial landscape

Predictor grain represented by coloured cells, response grain represented by black grid. 

```{r F1_methods_1}
# large landscape
ls_r <- nlm_random(42, 42) 
ls <- ls_r %>% as.data.frame(xy = TRUE)

# grid of study site at response grain
corners <- matrix(c(1,1,1,41,41,41,41,1,1,1), ncol = 2, byrow = TRUE)
ext <- st_polygon(list(corners))
grid <- st_make_grid(ext, cellsize = 10) %>% st_sf

# plot
p1 <- ggplot() + 
  geom_raster(data = ls, aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = grid, fill = NA, colour = "black", size = 2) + 
  geom_sf(data = grid[1,], fill = NA, colour = "red", size = 2) + 
  coord_sf(datum = NA) + 
  scale_fill_viridis_c(name = "") + 
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        axis.title = element_blank(), 
        panel.border = element_blank()) + 
  ggtitle("1. Define the response and predictor grains")

p1
```

## Moving window

For one cell of the landscape at the response grain

```{r F1_methods_2}
ls_sr <- crop(ls_r, st_sf(st_buffer(grid[1,], 1)))
ls_s <- ls_sr %>% as.data.frame(xy = TRUE)

w1 <- data.frame(x = c(0, 3, 3, 0, 0),
                 y = c(9, 9, 12, 12, 9))
w2 <- mutate(w1, x = x+1)
w3 <- data.frame(x = c(5, 8, 8, 5, 5),
                 y = c(5, 5, 8, 8, 5))
w4 <- data.frame(x = c(9, 12, 12, 9, 9),
                 y = c(0, 0, 3, 3, 0))

w1_plot <- ggplot() + 
  geom_raster(data = ls_s, aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = grid[1,], fill = NA, colour = "red", size = 2) + 
  coord_sf(datum = NA) + 
  geom_polygon(data = w1, aes(x =  x, y = y), fill = "grey", colour = "black", alpha = 0.7) + 
  annotate("text", x = 1.5, y = 10.5, 
           label = "italic(f(x))", parse = TRUE, size = 2.5) +
  annotate("text", x = 0, y = 13.2, hjust = 0, 
           label = "Predictor grain", size = 2.5) +
  annotate("text", x = 0, y = 14.6, hjust = 0, 
           label = "Scale of effect", size = 2.5) +
  annotate("text", x = 0, y = 16, hjust = 0, 
           label = "Response grain", size = 2.5) +
  geom_segment(aes(x=0, xend=1, y=12.5, yend=12.5), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both")) +
  geom_segment(aes(x=0, xend=3, y=13.9, yend=13.9), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both")) + 
  geom_segment(aes(x=0, xend=12, y=15.3, yend=15.3), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both")) + 
  scale_fill_viridis_c() + 
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        axis.title = element_blank(), 
        panel.border = element_blank(),
        legend.position = "none")

w2_plot <- ggplot() + 
  geom_raster(data = ls_s, aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = grid[1,], fill = NA, colour = "red", size = 2) + 
  coord_sf(datum = NA) + 
  geom_polygon(data = w2, aes(x =  x, y = y), fill = "grey", colour = "black", alpha = 0.7) + 
  annotate("text", x = 2.5, y = 10.5, 
           label = "italic(f(x))", parse = TRUE, size = 2.5) +
  annotate("text", x = 0, y = 13.2, hjust = 0, 
           label = "Predictor grain", size = 2.5, colour = "white") +
  annotate("text", x = 0, y = 14.6, hjust = 0, 
           label = "Scale of effect", size = 2.5, colour = "white") +
  annotate("text", x = 0, y = 16, hjust = 0, 
           label = "Response grain", size = 2.5, colour = "white") +
  geom_segment(aes(x=0, xend=1, y=12.5, yend=12.5), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both"), 
               colour = "white") +
  geom_segment(aes(x=0, xend=3, y=13.9, yend=13.5), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both"), 
               colour = "white") + 
  geom_segment(aes(x=0, xend=10, y=15.3, yend=14.5), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both"), 
               colour = "white") + 
  scale_fill_viridis_c() + 
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        axis.title = element_blank(), 
        panel.border = element_blank(),
        legend.position = "none")

w3_plot <- ggplot() + 
  geom_raster(data = ls_s, aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = grid[1,], fill = NA, colour = "red", size = 2) + 
  coord_sf(datum = NA) + 
  geom_polygon(data = w3, aes(x =  x, y = y), fill = "grey", colour = "black", alpha = 0.7) + 
  annotate("text", x = 6.5, y = 6.5, 
           label = "italic(f(x))", 
           parse = TRUE, 
           size = 2.5) +
  annotate("text", x = 0, y = 13.2, hjust = 0, 
           label = "Predictor grain", size = 2.5, colour = "white") +
  annotate("text", x = 0, y = 15.6, hjust = 0, 
           label = "Scale of effect", size = 2.5, colour = "white") +
  annotate("text", x = 0, y = 16, hjust = 0, 
           label = "Response grain", size = 2.5, colour = "white") +
  geom_segment(aes(x=0, xend=1, y=12.5, yend=12.5), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both"), 
               colour = "white") +
  geom_segment(aes(x=0, xend=3, y=13.9, yend=13.5), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both"), 
               colour = "white") + 
  geom_segment(aes(x=0, xend=10, y=15.3, yend=15.5), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both"), 
               colour = "white") + 
  scale_fill_viridis_c() + 
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        axis.title = element_blank(), 
        panel.border = element_blank(),
        legend.position = "none")

w4_plot <- ggplot() + 
  geom_raster(data = ls_s, aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = grid[1,], fill = NA, colour = "red", size = 2) + 
  coord_sf(datum = NA) + 
  geom_polygon(data = w4, aes(x =  x, y = y), fill = "grey", colour = "black", alpha = 0.7) + 
  annotate("text", x = 10.5, y = 1.5, 
           label = "italic(f(x))", 
           parse = TRUE, 
           size = 2.5) +
  annotate("text", x = 0, y = 13.2, hjust = 0, 
           label = "Predictor grain", size = 2.5, colour = "white") +
  annotate("text", x = 0, y = 15.6, hjust = 0, 
           label = "Scale of effect", size = 2.5, colour = "white") +
  annotate("text", x = 0, y = 16, hjust = 0, 
           label = "Response grain", size = 2.5, colour = "white") +
  geom_segment(aes(x=0, xend=1, y=12.5, yend=12.5), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both"), 
               colour = "white") +
  geom_segment(aes(x=0, xend=3, y=13.9, yend=13.5), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both"), 
               colour = "white") + 
  geom_segment(aes(x=0, xend=10, y=15.3, yend=15.5), size = 0.5,
               arrow = arrow(length = unit(0.1, "cm"), ends = "both"), 
               colour = "white") + 
  scale_fill_viridis_c() + 
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        axis.title = element_blank(), 
        panel.border = element_blank(),
        legend.position = "none")

#plot_grid(w1_plot, w2_plot, w3_plot, w4_plot, nrow = 2, axis = "b")

p2 <- w1_plot + ggtitle("2. Define the scale of effect and measure of variability,\n    and calculate for each cell at the predictor grain")

p2
```

## Moving window values

```{r F1_method_3}
ls_wm <- winmove(ls_r, d = 1, type = "rectangle", fun = "var") %>% as.data.frame(xy = TRUE)

p3 <- ggplot() + 
  geom_raster(data = ls_wm, aes(x = x, y = y, fill = layer)) + 
  geom_sf(data = grid, fill = NA, colour = "black", size = 2) + 
  geom_sf(data = grid[1,], fill = NA, colour = "red", size = 2) + 
  coord_sf(datum = NA) + 
  scale_fill_viridis_c(name = "") + 
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        axis.title = element_blank(), 
        panel.border = element_blank()) + 
  ggtitle("3. Calculate the predictor grain moving window values")

p3
```

# MWDA values

We take the mean for each response cell of the moving window variance to get a cell value. 

```{r F1_methods_4}
grid$ls_wma <- winmove_agg(grid, ls_r, 1, "rectangle", "var")

p4 <- ggplot() + 
  geom_sf(data = grid, aes(fill = ls_wma)) + 
  geom_sf(data = grid[1,], fill = NA, colour = "red", size = 2) + 
  coord_sf(datum = NA) + 
  scale_fill_viridis_c(name = "") + 
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        axis.title = element_blank(), 
        panel.border = element_blank()) + 
  ggtitle("4. Calculate the mean value of the moving window value for\n    each response grain cell")

p4
```

## Final plot for paper

```{r F1_methods, fig.width = 8.66142, fig.height = 8.66142}
plot_grid(p2, p3, p4, nrow = 1)
```

Graphical representation of the moving window data aggregation (MWDA) method. In calculating the MWDA measure, three aspects of scale are considered. Predictor grain is the characteristic spatial scale of the predictor variable, i.e. the resolution of the environmental data; scale of effect determines the appropriate scale of the relationship between predictor and response, for example an ecological neighbourhood; response grain is the grain of the unit into which you are predicting, i.e. the resolution of the response variable.  1. Define the scale of the response and predictor data (this is likely to be driven by the data sources). Predictor data shown in colour gradient, response data as black grid. 2. Define the scale-of-effect (neighbourhood size) and the measure of variability. 3. Calculate this for each cell at the predictor grain scale. 4. Aggregate the moving-window calculated measure to the response grain scale (in this case by taking a mean). In all four panels, the red outline represents the same spatial extent, allowing us to follow the calculation for this response grain cell.