---
title: "Appendix I: simulated landscapes"
output: pdf_document
header-includes:
    - \usepackage{booktabs}
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.path = 'figures/',
                      dpi = 300,
                      dev = c('pdf', 'tiff', 'png'),
                      fig.width = 6.496065,
                      fig.height = 6.496065) 
```

# Required packages

In the first step, we load all of the required packages. Note that grainchanger (the package developed for this paper) can be installed using:

`devtools::install_github("laurajanegraham/grainchanger")`

```{r setup}
# load required libraries
library(captioner)
library(knitr)
library(raster)
library(grainchanger)
library(broom)
library(landscapetools)
library(NLMR)
library(tidyverse)
library(patchwork)
library(cowplot)

# set up plotting options
theme_set(theme_bw(base_size = 7) + theme(strip.background = element_blank(), 
                             panel.grid.major = element_blank(),
                             panel.grid.minor = element_blank()))

# required for captioning and numbering tables and figures
tabs <- captioner(prefix = "Table AI.")
figs <- captioner(prefix = "Figure AI.")
```

# Simulated landscapes

1. Simulate study landscapes each with 100 coarse grid cells (each of 65 x 65 fine grid cells) with varying combinations of spatial autocorrelation using `nlm_fbm` from the `NLMR` package [@Sciaini2018]. This is the fractal Brownian motion method where `fract_dim` controls the level of spatial autocorrelation in the landscape. A value close to zero is a rough landscape (random), and a value of one is a smooth landscape (spatially autocorrelated). We used 5 combinations of spatial autocorrelation:

  - No spatial autocorrelation: `fractal_dim` = 0.1 for all landscapes
  - Low, varied spatial autocorrelation: `fractal_dim` in range 0.1--0.5
  - Varied spatial autocorrelation: `fractal_dim` in range 0.1--1
  - High, varied spatial autocorrelation: `fractal_dim` in range 0.5--1
  - High spatial autocorrelation: `fractal_dim` = 1 for all landscapes
  
2. Create categorical landscapes with 5 land cover classes by using `util_classify` with random weights for each class. 

3. Use `winmove` from the `grainchanger` package to calculate the moving window measures (MWM) for each coarse grid cell. For continuous landscapes, this is variance; for categorical landscapes, this is Shannon evenness. We do this for 5 window sizes:

  - 3 x 3 pixels (0.2% of coarse grid cell)
  - 9 x 9 pixels (1.9% of coarse grid cell)
  - 15 x 15 pixels (5.3% of coarse grid cell)
  - 35 x 35 pixels (29% of coarse grid cell)
  - 55 x 55 pixels (71.6% of coarse grid cell)
  
4. Calculate the corresponding measures for each coarse grid cell using standard (non-moving window) methods (LSM).

We created 100 replicate study landscapes using the Iridis HPC at University of Southampton. The below code is for one replicate:

```{r sim_ls, eval = FALSE}
library(raster)
library(NLMR)
library(landscapetools)
library(grainchanger)
library(tidyverse)
library(furrr)

plan(multiprocess)

# 1. Create the landscapes and calculate LS and MW measures 
# Set up spatial autocorrelation parameters

strt <- Sys.time()
res <- bind_rows(
  tibble(sa_scenario = "No spatial autocorrelation", sa_values = rep(0.1, 100)),
  tibble(sa_scenario = "Low, varied spatial autocorrelation", sa_values = seq(0.1, 0.5, length.out = 100)),
  tibble(sa_scenario = "Varied spatial autocorrelation", sa_values = seq(0.1, 1, length.out = 100)),
  tibble(sa_scenario = "High, varied spatial autocorrelation", sa_values = seq(0.5, 1, length.out = 100)),
  tibble(sa_scenario = "High spatial autocorrelation", sa_values = rep(1, 100))
) %>% 
  # create the continuous and categorical landscapes
  mutate(cont_ls = future_map(sa_values, function(x) nlm_fbm(ncol = 400, nrow = 400, resolution = 25, fract_dim = x)),
         # the categorical map has a random proportion between landscapes for each landscape
         cat_wt = future_map(sa_values, function(x) diff(c(0, sort(runif(4)), 1))),
         cat_ls = future_map2(cont_ls, cat_wt, function(x, wt) util_classify(x, weighting = wt)),
         # calculate LSM
         cont_lsm = future_map_dbl(cont_ls, function(x) x %>% raster::values() %>% var),
         cat_lsm = future_map_dbl(cat_ls, function(x) diversity(x, lc_class = 0:4))) %>% 
  crossing(window = c(500, 1000, 1500, 3500)) %>% 
  mutate(cont_ls_pad = future_map2(cont_ls, window, create_torus),
         cat_ls_pad = future_map2(cat_ls, window, create_torus),
         # calculate MWM
         cont_mwm = future_map2_dbl(cont_ls_pad, window, function(ls, w) winmove(ls, d = w, type = "rectangle", fun = "var") %>% trim %>% raster::values() %>% mean),
         cat_mwm = future_map2_dbl(cat_ls_pad, window, function(ls, w) winmove(ls, d = w, type = "rectangle", fun = "shei", lc_class = 0:4) %>% trim %>% raster::values() %>% mean)
  ) %>% select(-cont_ls, -cat_ls, -cont_ls_pad, -cat_ls_pad)

runtime = difftime(Sys.time(),  strt)
out = list(res = res, runtime = runtime)
save(out, file = tempfile(tmpdir = ".", fileext = ".Rda")) 
```

We use the simulated landscapes to explore two questions:

1. What is the correlation between the moving window and standard aggregations at different levels of spatial autocorrelation and moving window sizes? 

2. Can we successfully identify the scale of effect?

# Question 1: correlation between MW and LS

Here we calculate the Spearman's rank correlation between LSM and MWM for each replicate at each level of spatial autocorrelation and moving window size. The plot below shows the mean and standard deviation of these correlations (across 100 replicates). 

```{r load_sims}
load("results/sims_all_replicates.Rda")

df <- df %>% 
  mutate(sa_scenario = factor(sa_scenario, 
                              levels = c("No spatial autocorrelation", 
                                         "Low, varied spatial autocorrelation", 
                                         "Varied spatial autocorrelation",
                                         "High, varied spatial autocorrelation",
                                         "High spatial autocorrelation")),
         window = factor(window, labels = unique(paste0(round(((2*window + 1)^2/10000^2)*100, 1), "%"))),
         rep = ceiling((rep %>% as.factor %>% as.numeric)/10)) %>% 
  group_by(rep, sa_scenario, window) %>% 
  mutate(ls_id = 1:n()) %>% 
  ungroup()
```

```{r correlations}
corrs <- df %>% 
  group_by(rep, sa_scenario, window) %>% 
  nest() %>% 
  mutate(
    cont_cor = map_dbl(data, function(x) {
    cor.test(x$cont_lsm, x$cont_mwm, method = "spearman")$estimate
  }),
  cat_cor = map_dbl(data, function(x) {
    cor.test(x$cat_lsm, x$cat_mwm, method = "spearman")$estimate}
  )) %>% 
  select(-data) %>% 
  group_by(sa_scenario, window) %>% 
  summarise_at(.vars = vars(cont_cor, cat_cor), .funs = c("mean", "sd")) %>% 
  gather(measure, value, -sa_scenario, -window) %>% 
  mutate(measure = factor(measure, 
                          labels = c("Categorical_mean", "Categorical_sd", 
                                     "Continuous_mean", "Continuous_sd"))) %>% 
  separate(measure, into = c("ls_type", "measure_type"), sep = "_")

# 1. correlations between MW and LS for diff spatial autocorrelation and window sizes
p1 <- ggplot(corrs %>% filter(measure_type == "mean"), 
       aes(x = window, y = sa_scenario, fill = value)) + 
  geom_tile() + 
  scale_fill_distiller("MWDA vs SDA cor\n(mean)", palette = "BrBG", limits = c(-1, 1), direction = 1) + 
  labs(x = "Window % of landscape", y = "") + 
  facet_wrap(~ls_type)

p2 <- ggplot(corrs %>% filter(measure_type == "sd"), 
             aes(x = window, y = sa_scenario, fill = value)) + 
  geom_tile() + 
  scale_fill_distiller("MWDA vs SDA cor\n(SD)", palette = "BrBG") + 
  labs(x = "Window % of landscape", y = "") + 
  facet_wrap(~ls_type)


p1 / p2
```

- Higher correlation between MWM and LSM in continuous landscapes
- We even get negative correlation in categorical landscapes for small moving window and variation in spatial autocorrelation. 
- MW and LS are highly correlated for the two largest window sizes. 
- There is more variability between the amount of correlation in replicates for continuous than categorical landscapes. 

Why do we get negative correlations in some cases? The below plots show each spatial autocorrelation scenario and window size combination which results in a negative correlation between MWM and LSM measures. In the first plot, points are coloured by the spatial autocorrelation value used to generate the landscape. In the second, they are coloured by the replicate ID.    

```{r}
neg_corr <- corrs %>% filter(value < 0) %>% 
  inner_join(df)

neg_corr_p1 <- ggplot(neg_corr, aes(x = cont_lsm, y = cont_mwm, colour = sa_values, group = sa_values)) +
  geom_point() + 
  facet_wrap(~ sa_scenario + window)

neg_corr_p2 <- ggplot(neg_corr, aes(x = cont_lsm, y = cont_mwm, colour = rep, group = rep)) +
  geom_point() +
  facet_wrap(~ sa_scenario + window)

neg_corr_p1 + neg_corr_p2
```

In all cases, low values of spatial autocorrelation result in high MWM and low LSM variance; high values of spatial autocorrelation result in low MWM and high LSM. This is what is driving the negative correlation. Within spatial autocorrelation values, we still see positive correlation between replicates, these correlations are greater for lower values of spatial autocorrelation. Within replicate, this manifests as negative correlation (leading to an overall negative correlation when taken as the mean of all replicates). 

## Question 2: identifying the scale-of-effect

In order to examine whether we can correctly identify the scale of effect when comparing several correlated measures calculated using different window sizes, we use our MW measures to generate an ecological response variable for every window size in each coarse grid cell in each replicate landscape and all spatial autocorrelation scenarios. We calculate this response variable as $y_w = MW_w + \epsilon$ where $y_w$ is the response variable for a given window size $w$, $MW_w$ is the moving window calculated measure for window size $w$ and $\epsilon ~ N(0, \sigma)$. We use three levels of $\epsilon$: 

- low $\epsilon$: first percentile of the MWM measure value within each spatial autocorrelation and window combination. This represents data with minimal noise. 
- moderate $\epsilon$: tenth percentile of the MWM measure value within each spatial autocorrelation and window combination. This represents data with a moderate amount of noise. 
- high $\epsilon$: median of the MWM measure value within each spatial autocorrelation and window combination. This represents data with a large amount of noise. 

For each $y_w$ we fit a univariate linear model with each $MW_w$ as the covariate and use $R^2$ and AIC to examine whether the correct window size (and thus, scale of effect) is identified. 

For the models fit for each of the $y_w$ and $MW_w$ combinations, we calculate the number of replicates where the best fitting model was the correct one. 

```{r scale_check_low, fig.width = 6, fig.height = 6}
window_y = unique(df$window)
noise_level = tibble(noise_level = c("Low noise", "Moderate noise", "High noise"), 
                     noise_value = c(0.1, 0.25, 0.5))

df_narrow <- df %>% select(rep, ls_id, sa_scenario, window, cont_mwm, cat_mwm) %>% 
  gather(measure, value, -rep, -ls_id, -sa_scenario, -window)

y_vals <- df_narrow %>% 
  crossing(noise_level) %>% 
  group_by(sa_scenario, window, measure) %>% 
  mutate(sigma = noise_value * mean(value),
         eps = map_dbl(sigma, function(x) rnorm(1, 0, x)),
         y = value + eps) %>% 
  ungroup() %>% 
  select(rep, sa_scenario, ls_id, window_y = window, noise_level, measure, y)

mod_df <- df_narrow %>% rename(window_x = window) %>% 
  crossing(window_y) %>% 
  inner_join(y_vals) %>% 
  group_by(rep, sa_scenario, noise_level, measure, window_x, window_y) %>% 
  nest() %>% 
  mutate(mod = map(data, function(x) lm(y ~ value, data = x)),
         mod_glance = map(mod, glance),
         mod_r2 = map_dbl(mod_glance, function(x) x$r.squared),
         mod_aic = map_dbl(mod_glance, function(x) x$AIC))

correct_df <- mod_df %>% 
  select(rep, sa_scenario, window_x, window_y, noise_level, measure, mod_r2, mod_aic) %>% 
  gather(key, value, -rep, -sa_scenario, -window_x, -window_y, -noise_level, -measure) %>% 
  mutate(value = case_when(key == "mod_r2" ~ 1 - value, TRUE ~ value),
         ls_type = factor(measure, labels = c("Categorical", "Continuous")),
         fit_measure = factor(key, 
                               labels = c("AIC", expression(R^{2}))),
         noise_level = factor(noise_level, 
                              levels = c("Low noise", "Moderate noise", "High noise"))) %>% 
  group_by(rep, sa_scenario, window_y, noise_level, ls_type, fit_measure) %>% 
  slice(which.min(value)) %>% 
  mutate(pass = window_x == window_y) %>% 
  group_by(sa_scenario, window_y, noise_level, ls_type, fit_measure) %>% 
  summarise(pass = sum(pass)) %>% 
  ungroup()

scale_plot <- ggplot(correct_df %>% filter(fit_measure == "AIC"), 
       aes(x = window_y, y = sa_scenario, fill = pass)) + 
  geom_tile() + 
  scale_fill_viridis_c("% correct") + 
  labs(x = "Window % of landscape", y = "") + 
  facet_grid(noise_level ~ ls_type)

p1/scale_plot + plot_layout(heights = c(1,3)) + plot_annotation(tag_levels = "a", tag_suffix = ")")

correct_df %>% filter(fit_measure == "AIC") %>% 
  group_by(ls_type, noise_level) %>% 
  summarise_at(.vars = "pass", .funs = c("mean", "min", "max"))

correct_df %>% filter(fit_measure == "AIC") %>% 
  group_by(ls_type, noise_level) %>% 
  slice(which.min(pass))
```

- Whether judged by $R^2$ or AIC, we are able to identify the correct scale of effect in low noise datasets most of the time
- In moderate and high noise datasets, we are best able to identify the correct scale of effect when it is small or large, and the data is continuous. 
- Overall, we were least successful in identifying the scale of effect in the No spatial autocorrelation scenario

It's also useful to understand the general fit of the models at each spatial autocorrelation and window size combination. 

```{r, fig.width = 6, fig.height = 6}
r2_df <- mod_df %>% 
  select(measure, noise_level, sa_scenario, window_x, window_y, mod_r2) %>% 
  filter(window_x == window_y) %>% 
  mutate(ls_type = factor(measure, labels = c("Categorical", "Continuous")),
         noise_level = factor(noise_level, 
                              levels = c("Low noise", "Moderate noise", "High noise"))) %>% 
  group_by(ls_type, noise_level, sa_scenario, window_x) %>% 
  summarise(mean_r2 = round(mean(mod_r2), 2))

ggplot(r2_df, aes(x = window_x, y = sa_scenario, fill = mean_r2)) + 
  geom_tile() + 
  scale_fill_viridis_c(expression(R^2)) + 
  labs(x = "Window % of landscape", y = "") + 
  facet_grid(noise_level ~ ls_type)

r2_df %>% group_by(ls_type, noise_level) %>% 
  summarise_at(.vars = "mean_r2", .funs = c("min", "mean", "max"))
```

# Session Info

```{r session_info}
session <- devtools::session_info()
session[[1]]
session[[2]] %>% kable
```